---
title: "AN588-Week-2-cmetcalf"
author: "Cat Metcalf"
date: "2023-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# AN588-Week-2 R Markdown HW1



{stringr} did not work initially --> go to Tools>Install Packages...>CRAN selected>type package name OR install.packages("stringr")


```{r m <- matrix(data = c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3, byrow = FALSE)
m}
*copy and paste broke the chunk ERROR*

*chunk error again* >> keep out of brackets and soemtimes have to retype the "}"

**L: **I don't quite understand the error happening here and your method to fix this. Maybe you can try explaining the problem you encountered more clearly? 

# Challenge 1:

```{r}
v <- c("Stately", "plump", "Buck", "Mulligan", "came", "from", "the", "stairhead", "bearing", "a", "bowl", "of", "lather", "on", "which", "a", "mirror", "and", "a", "razor", "lay", "crossed")
v
```

```{r}
strsplit(x = v, split = 3,fixed = FALSE, perl = FALSE, useBytes = FALSE)
```

```{r}
v[seq(from = 1, to = 22, by = 3)]
```
*success of every 3rd element* 

**L: **Nice way to do it! I feel like the sequencing should start from 3 instead of 1, since it will actually give you the result of the 1^st^ and 4^th^ word, etc. 

```{r}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```

**L: **I am not quite sure but maybe the matrix assigning above is part of the content of Challenge 2? If that's the case, maybe you will want to move it to the place of challenge 2 below. 

# Challenge 2:

Solution below from: <https://stats.stackexchange.com/questions/10225/extracting-multiple-columns-from-a-matrix-in-r>

```{r}
x <- m [,c(2,3,6)]
x
```
 *the comma outside the "c" command indicates whether it is for row or column extraction aka [row, column] 

**L: **Good explanation!:)
 
```{r}
x <- m [c(6, 7, 8),]
x
```

Given the matrix, m, above, extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x:

```{r}
x <- m[2:6, 2:9]
x
```
*complete success!*

**L: **I like how you cite the resource you looked at when doing the homework! Everything looks good for this challenge. :)

# Challenge 3:

Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1 (i.e., a descending series)


```{r}
a <- array(data= 400:1, dim= c(5,5,4,4))
a
```

Dimensions go in the commas after "c".

```{r}
a[1,1,1,2]
```

```{r}
a[2,3,2,]
```

```{r}
a[1:5, 1:5, 3, 3]
```

```{r}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```

```{r}
m[7, 1]
```


s <- c("this", "is", "a", "vector", "of", "strings") 
*here we have a vector setup*

# Challenge 4
```{r}
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")
Strepsirhini
```

```{r}
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Anthropoidea <- list(Platyrrhini, Catarrhini)
names(Anthropoidea) <- list("Platyrrhini", "Catarrhini")
Tarsioidea <- c("Tarsiidae")
Haplorhini <- list(Anthropoidea, Tarsioidea)
names(Haplorhini) <- list("Anthropoidea", "Tarsioidea")
```

Each time you list it will put it below whatever you tell it to list TO.

```{r}
Haplorhini
```

```{r}
df <- read.csv("~/Desktop/random-people.csv", sep = ",", header = TRUE, stringsAsFactors = FALSE)
df[, c(1, 3, 4, 11, 12)]
```

You have to have the file in the location stated. (Ex.Desktop)

```{r}
str(df)
```

```{r}
df[, 3]
```
```{r}
str(df[, 3])
```
returns a vector of data in column 3
```{r}
str(df$name.last)
```

```{r}
new_df <- df[df$school == "BU", ]
new_df
```
(shows 0 because skipped parts where you add this in module steps)

# Challenge 5:
```{r}
m3 <- matrix(data = c(3, 0, 1, 23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow=5, ncol=3, byrow=TRUE)
m3
```

byrow must be true for this to fill by row!

```{r}
data.frame(m3)
```

```{r}
 df[,2]<-as.character(df[,2])
str(df)
```

```{r}
 df[,3]<-as.factor(df[,3])
str(df)
```
Gives a null? -> as.factor ; have to specify data frame and what you are changing to a factor
```{r}
str(m3)
```

*Module 4 Complete*