---
title: "cmetcalf_FinalHomeworkCode_01"
author: "Cat Metcalf"
date: "2023-09-18"
output: tufte::tufte_html
---

# AN588-Week-2 R Markdown HW1

***

### Downloading stringr
{stringr} did not work initially --> go to Tools>Install Packages...>CRAN selected>type package name OR install.packages("stringr")

### Learning Chunks
```{r m <- matrix(data = c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3, byrow = FALSE)
m}

Keep code out of brackets and soemtimes have to retype the "}" after the r in the chunk, or else the play button disappears!

*** 

## Challenge 1:

```{r creating a word vector}
v <- c("Stately", "plump", "Buck", "Mulligan", "came", "from", "the", "stairhead", "bearing", "a", "bowl", "of", "lather", "on", "which", "a", "mirror", "and", "a", "razor", "lay", "crossed")
v
```

```{r using strsplit}
strsplit(x = v, split = 3,fixed = FALSE, perl = FALSE, useBytes = FALSE)
```

```{r extracting every 3rd word}
v[seq(from = 3, to = 22, by = 3)]
```
*success of every 3rd element, which should start at 3!* 

## Challenge 2
```{r creating a matrix}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```
 
Solution below from: <https://stats.stackexchange.com/questions/10225/extracting-multiple-columns-from-a-matrix-in-r>

```{r extracting columns of data}
x <- m [,c(2,3,6)]
x
```
 *the comma outside the "c" command indicates whether it is for row or column extraction aka [row, column] 
 
```{r extracting rows of data}
x <- m [c(6, 7, 8),]
x
```

Given the matrix, m, above, extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x:

```{r extracting rows and columns in a section}
x <- m[2:6, 2:9]
x
```
*complete success!*

## Challenge 3:

Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1 (i.e., a descending series)


```{r creating an array}
a <- array(data= 400:1, dim= c(5,5,4,4))
a
```

Dimensions go in the commas after "c".

```{r extracting values in array}
a[1,1,1,2]
```

```{r extracting values in array 2}
a[2,3,2,]
```

```{r extracting values in array 3}
a[1:5, 1:5, 3, 3]
```

```{r creating a matrix again}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```

```{r extracting one specific value}
m[7, 1]
```


## Challenge 4
```{r learning to do Primate lists}
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")
Strepsirhini
```

```{r applying Primate listing}
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Anthropoidea <- list(Platyrrhini, Catarrhini)
names(Anthropoidea) <- list("Platyrrhini", "Catarrhini")
Tarsioidea <- c("Tarsiidae")
Haplorhini <- list(Anthropoidea, Tarsioidea)
names(Haplorhini) <- list("Anthropoidea", "Tarsioidea")
```

Each time you list it will put it below whatever you tell it to list TO.

```{r Primate listing result}
Haplorhini
```

```{r file and data retrieval}
df <- read.csv("~/Desktop/random-people.csv", sep = ",", header = TRUE, stringsAsFactors = FALSE)
df[, c(1, 3, 4, 11, 12)]
```

You have to have the file in the location stated. (Ex.Desktop) 
**Important note**: (referenced from Chuhan) Using getwd() and setwd() allow you to change default route and only type file name to retrieve it.

```{r str function with data frame}
str(df)
```

```{r extraction in data frame}
df[, 3]
```

```{r further structuring for data and extraction in a column}
str(df[, 3])
```
returns a vector of data in column 3
```{r structure by last name}
str(df$name.last)
```

```{r new dataframe based on school}
new_df <- df[df$school == "BU", ]
new_df
```
(shows 0 because skipped parts where you add this in module steps, please reference module parts for full steps and answer, this is to show how to sort)

## Challenge 5:
```{r filling matrix by row}
m3 <- matrix(data = c(3, 0, 1, 23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow=5, ncol=3, byrow=TRUE)
m3
```

byrow must be true for this to fill by row!

```{r assigning new data frame}
df1 <- data.frame(m3)
```
Have to assign new dataframe to "df1" above in order for the below chunk to work and output the correct numbers, as with regular df it outputs from previous data frame instead of the correct one! (Peer Commentor Chuhan assisted in fixing this error)

```{r coercing to logical-valued}
 df1[,2]<-as.character(df1[,2])
str(df1)
```

```{r coercing to factor-valued}
 df1[,3]<-as.factor(df1[,3])
str(df1)
```
Gives a null? -> as.factor ; have to specify data frame and what you are changing to a factor
```{r structure new data frame}
str(m3)
```

*Module 4 Complete*

***

## Reflection:

I came across a difficulty in making things logical valued and factor valued in the later challenge, mainly because I was not being specific about what data frame the code was being applied to. Another difficulty I came across at first was extracting values from a matrix since I did not make the connection that the brackets delineated the specific rows and columns in the order [row, columns] since I would have thought it was [columns, row]. Finally, it is a struggle that github does not allow you to easily clone a file but lets you do it to a repo, although copy and paste does work wonders.

The commentary was extremely helpful, first in helping me troubleshoot and fix things that went wrong with my code. For example, I had trouble with the data frame in the final parts of Challenge 5 because the outputs were giving answers from my previous data frame, and my peer commentor gave a suggestion that made me able to 1) understand my mistake and 2) figure out how to fix it to aid me with future data frame structuring. Second, the commentary allowed me to feel good about certain parts of my code that I may have felt were weaker, since my commentor not only gave constructive criticism but also uplifting remarks like saying I did a good job! Lastly, being able to see how other people solved problems let me reflect on my own code and how I could make my code cleaner.

I enjoyed playing with the "prettification" aspect of this assignment since it let me learn how to make code easily accessible to other people. It was also nice to learn how to extract values from datasets, as I see how that can be very helpful for future research when you are short on time and need to only apply things to certain valuables. One thing to be improved is the explanation for cloning a repository and the files, as it would really benefit to have that in a concrete document to reference for future projects since it was not as simple as pressing a button or two (Unless it is?!). Another improvement would be differentiation between actual challenge parts and instructive parts, as even my peer commentor noted in their comments on my code that this made it confusing for what to include in our file.